---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

# khroma

[![Build Status](https://travis-ci.org/nfrerebeau/khroma.svg?branch=master)](https://travis-ci.org/nfrerebeau/khroma)
[![CRAN status](https://www.r-pkg.org/badges/version/khroma)](https://cran.r-project.org/package=khroma)
[![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)

## Overview

khroma is a utility package: it provides usefull colour schemes for archaeological data visualization. khroma provides Paul Tol's colour schemes and several thematic palettes (geologic timescale, etc.) with scales for ggplot2.

## Installation

Install the released version of khroma from CRAN:

```{r cran-installation, eval = FALSE}
install.packages("khroma")
```

Or install the development version from GitHub with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("nfrerebeau/khroma")
```

## Usage
### Paul Tol's Colour Schemes

Paul Tol offers carefully chosen schemes, ready for each type of data, with colours that are:

* distinct for all people, including colour-blind readers;
* distinct from black and white;
* distinct on screen and paper;
* matching well together.

All the scales presented in his 2018 technical note are implemented here, for use with graphics or ggplot2.

#### Qualitative data

```{r tol-qualitative, echo=FALSE, fig.height=7, fig.width=7}
# Qualitative data
par(mfrow = c(2, 2), mar = c(0,0,0,0), pty = "s")
scales::show_col(khroma::colour("bright")(7))
scales::show_col(khroma::colour("vibrant")(7))
scales::show_col(khroma::colour("muted")(9))
scales::show_col(khroma::colour("light")(9))
```

#### Diverging data

```{r tol-diverging, echo=FALSE, fig.height=7, fig.width=7}
# Diverging data
par(mfrow = c(2, 2))
scales::show_col(khroma::colour("BuRd")(9))
scales::show_col(khroma::colour("PRGn")(9))
scales::show_col(khroma::colour("sunset")(11),
                 cex_label = 0.7)
```

#### Sequential data

```{r tol-sequential, echo=FALSE, fig.height=3.5, fig.width=7}
# Sequential data
par(mfrow = c(1, 2))
scales::show_col(khroma::colour("YlOrBr")(9))
scales::show_col(khroma::colour("smooth rainbow")(34),
                 cex_label = 0.7)
```

### Scientific colour schemes

For now, the only scientific colour scheme available is that of the International Chronostratigraphic Chart. More will be added in future releases (suggestions are welcome).
