---
title: "Paul Tol's Colour Schemes"
author: "N. Frerebeau"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    number_sections: yes
    fig_caption: yes
    toc: true
bibliography: bibliography.bib
vignette: >
  %\VignetteIndexEntry{Paul Tol's Colour Schemes}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
 fig.align = "center"
)
```

```{r setup}
library(khroma)
```

# Introduction
@tol2018 offers carefully chosen schemes, ready for each type of data, with colours that are:

* Distinct for all people, including colour-blind readers.
* Distinct from black and white.
* Distinct on screen and paper.
* Matching well together.

All the scales presented in Paul Tol's technical note are implemented here, for use with base R `graphics` or `ggplot2`.

# Qualitative data
According to Paul Tol's technical note, the *bright*, *contrast*, *vibrant* and *muted* colour schemes are colour-blind safe.

The *light* colour scheme is reasonably distinct for both normal or colour-blind vision and is intended to fill labelled cells.

The *pale* and *dark* schemes are not very distinct in either normal or colour-blind vision and should be used as a text background or to highlight a cell in a table.

The qualitative colour schemes must be used as given (no interpolation): colours are picked up to the maximum number of supported values. Refer to the original document for details about the recommended uses (see references).

| Scheme   | Max. colours |
|:---------|-------------:|
| bright   | 7            |
| contrast | 3            |
| vibrant  | 7            |
| muted    | 9            |
| pale     | 6            |
| dark     | 6            |
| light    | 9            |

## *bright*
```{r tol_quali_bright, fig.height = 2, fig.width = 7}
bright <- colour("bright")

plot_scheme(bright(7), colours = TRUE, names = TRUE, size = 0.9)
plot_map(bright(7))
```

## *contrast*
```{r tol_quali_contrast, fig.height = 2, fig.width = 7}
contrast <- colour("contrast")

plot_scheme(contrast(3), colours = TRUE, names = TRUE, size = 0.9)
plot_map(contrast(3))
```

## *vibrant*
```{r tol_quali_vibrant, fig.height=2, fig.width=7}
vibrant <- colour("vibrant")

plot_scheme(vibrant(7), colours = TRUE, names = TRUE, size = 0.9)
plot_map(vibrant(7))
```

## *muted*
```{r tol_quali_muted, fig.height=2, fig.width=7}
muted <- colour("muted")

plot_scheme(muted(9), colours = TRUE, names = TRUE, size = 0.9)
plot_map(muted(9))
```

## *pale* and *dark*
```{r tol_quali_pale_dark, fig.height=2, fig.width=7}
pale <- colour("pale")
dark <- colour("dark")

plot_scheme(pale(6), colours = TRUE, names = TRUE, size = 0.9)
plot_scheme(dark(6), colours = TRUE, names = TRUE, size = 0.9)
```

## *light*
```{r tol_quali_light, fig.height=2, fig.width=7}
light <- colour("light")

plot_scheme(light(9), colours = TRUE, names = TRUE, size = 0.9)
plot_map(light(9))
```

# Diverging data
## *sunset*
```{r tol_div_sunset, fig.height=2, fig.width=7}
sunset <- colour("sunset")

plot_scheme(sunset(9), colours = TRUE, size = 0.9)
plot_map(sunset(9))
```

## *BuRd*
```{r tol_div_BuRd, fig.height=2, fig.width=7}
BuRd <- colour("BuRd")

plot_scheme(BuRd(9), colours = TRUE, size = 0.9)
plot_map(BuRd(9))
```

## *PRGn*
```{r tol_div_PRGn, fig.height=2, fig.width=7}
PRGn <- colour("PRGn")

plot_scheme(PRGn(9), colours = TRUE, size = 0.9)
plot_map(PRGn(9))
```

# Sequential data
## *YlOrBr*
```{r tol_seq_YlOrBr, fig.height=2, fig.width=7}
YlOrBr <- colour("YlOrBr")

plot_scheme(YlOrBr(9), colours = TRUE, size = 0.9)
plot_map(YlOrBr(9))
```

## *iridescent*
```{r tol_seq_iridescent, fig.height=2, fig.width=7}
iridescent <- colour("iridescent")

plot_scheme(iridescent(9), colours = TRUE, size = 0.9)
plot_map(iridescent(9))
```

## *rainbow*
As a general rule, ordered data should not be represented using a rainbow scheme. There are three main arguments against such use [@tol2018]:

* The spectral order of visible light carries no inherent magnitude message.
* Some bands of almost constant hue with sharp transitions between them, can be perceived as jumps in the data.
* Colour-blind people have difficulty distinguishing some colours of the rainbow.

If such use cannot be avoided, Paul Tol's technical note provides two colour schemes that are reasonably clear in colour-blind vision. To remain colour-blind safe, these two schemes must comply with the following conditions:

*discrete rainbow*
: This scheme must not be interpolated.

*smooth rainbow*
: This scheme does not have to be used over the full range.

```{r tol_seq_rainbow, fig.height=2, fig.width=7}
discrete_rainbow <- colour("discrete rainbow")

plot_scheme(discrete_rainbow(12), colours = TRUE, size = 0.7)
plot_map(discrete_rainbow(12))
```

# References
